\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-07-26 paper/paper.tex
Updates.

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone is a replicated log plug-in for
key value stores that systematically addresses various sources of this
bottleneck. It uses a small amount of non-volatile memory directly addressable
by the CPU - such as in the form of NVDIMMs - to remove block oriented IO
devices such as SSDs from the critical path for appending to the log. It then
addresses network overheads using an implementation of the RAFT consensus
protocol that is designed around a userspace network stack to relieve the CPU of
the burden of data copies. Finally, it provides a way to efficiently map the
commutativity in key-value store APIs to the parallelism available in commodity
NICs via multiple transmit receive queues to scale replication performance on
multicore machines. Cyclone is able to replicate millions of small updates a
second using only commodity 10 gigabit ethernet adapters. As a practical
application, we use it to improve the performance and availability of RocksDB, a
popular persistent key value store by an order of magnitude when compared to its
own write ahead log without replication.
\end{abstract}

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. We describe Cyclone's
system architecture in Section~\ref{sec:sysarch}. Cyclone provides the order of
magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the host
- such as in the form of NVDIMMs~\cite{farm} or directly attached persistent
memory~\cite{pmfs}. Cyclone logs update operations to the key value store and
calls on the key value store to execute the update only when it has been
successfully replicated to majority quorum of replicas. All reads and write to
the key value store are always executed by the leader replica. A client side
stub provided by Cyclone automatically locates and directs read and write
requests to the leader replica, ensuring linearizable semantics from the key
value store. The novelty in Cyclone is that it accomplishes this without
compromising on the performance baseline of no logging or replication. Cyclone
does this by systematically addressing the sources of performance degradation in
Figure~\ref{fig:problem}.

The primary cause of slowdown when turning on a write ahead log on a single
machine is the need to do synchronous block IO for every update. For even
moderately sized updates, this leads to severe write amplification as a 4KB page
must be synchronously written to the log file on flash. Cyclone avoids this by
making use of directly attached persistent memory on the server. This can be in
the form of NVDIMMs (DIMMs with an added ultracapacitor to dump state to a small
amount of attached flash) or newer forms of persistent memory~\cite{pmfs}. For
current deployments, NVDIMMs are expensive - retailing at approximately 2X the
price of a standard DIMM [find ref]. This means that they will likely be
available in small quantities. We compensate for this by periodically draining
the NVDIMM log into a log placed on a standard SSD on the system. We describe
this two level log structure in Section~\ref{sec:storage}.

The network overheads for replicating a log entry to a quorum of machines stems
from the overhead of the TCP/IP stack in the linux kernel - itself the topic of
much research [dataplane OS stuff]. We therefore replase the TCP/IP stack with
DPDK, a userspace network library that allows direct IO of raw ethernet frames
to and from the network interface card (NIC) eliminating all overheads. We no
longer use TCP - since the rare network errors encountered in a typical
datacenter network are automatically handled by the replication protocol (RAFT)
as part of its failure model. Finally, we carefully design our implementation of
RAFT to relieve the CPU of all data copy operations, leaving the CPU to only
run the protocol and return an accept/reject decision for each received
packet. The set of optimizations to the network stack is decribed in
Section~\ref{sec:network}.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. To exploit
this parallelism we use Cyclone to replicate multiple logs, each using a
different instance of the RAFT consensus protocol. We show how to exploit the
commutativity inherent in key-value store interfaces to make use of this
parallelism in Cyclone's design in Section~\ref{sec:parallelism}.

A detailed evaluation of Cyclone's replication is provided in the context of
RocksDB - a popular persistent key value store - in
Section~\ref{sec:evaluation}. We show that Cyclone achieves performance
transparent replication in that it can provide both a persistent write ahead log
and replication without compromising on RocksDB's performance.

>>>>

\begin{abstract}
\vspace{0.1in}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone is a replicated log plug-in for
key value stores that systematically addresses various sources of this
bottleneck. It uses a small amount of non-volatile memory directly addressable
by the CPU - such as in the form of NVDIMMs - to remove block oriented IO
devices such as SSDs from the critical path for appending to the log. It then
addresses network overheads using an implementation of the RAFT consensus
protocol that is designed around a userspace network stack to relieve the CPU of
the burden of data copies. Finally, it provides a way to efficiently map the
commutativity in key-value store APIs to the parallelism available in commodity
NICs via multiple transmit receive queues to scale replication performance on
multicore machines. Cyclone is able to replicate millions of small updates a
second using only commodity 10 gigabit ethernet adapters. As a practical
application, we use it to improve the performance and availability of RocksDB, a
popular persistent key value store by an order of magnitude when compared to its
own write ahead log without replication.
\end{abstract}

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone entirely
closes the performance gap due to both the storage and the network shown in
Figure~\ref{fig:problem}. We show that Cyclone achieves performance transparent
replication in that it can provide both a persistent write ahead log and
replication without compromising on RocksDB's performance.

The first key contribution of Cyclone is in showing how a small amount of
non-volatile memory can be combined with the flash SSD based persistent logs
when providing high availability.  The primary cause of slowdown when turning on
a write ahead log on a single machine in Figure~\ref{fig:problem} is the need to
do synchronous block IO for every update. For evenmoderately sized updates, this
leads to severe write amplification as a 4KB page must be synchronously written
to the log file on flash. Cyclone avoids this by making use of directly attached
persistent memory on the server. This can be in the form of NVDIMMs (DIMMs with
an added ultracapacitor to dump state to a small amount of attached flash) or
newer forms of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs
are expensive - retailing at approximately 2X the price of a standard DIMM [find
  ref]. This means that they will likely be available in small quantities. We
compensate for this by periodically draining the NVDIMM log into a log placed on
a standard SSD on the system.

The second key contribution of this work is in showing that the performance
overheads in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum - a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory eliminating block IO from the persistence step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

The rest of the paper is organized as follows. We describe Cyclone's system
architecture in Section~\ref{sec:sysarch}. We describe this two level log
structure in Section~\ref{sec:storage}. The set of optimizations to the network
stack is decribed in Section~\ref{sec:network}. Both the storage (NVM, SSD) as
well as the NIC expose significant parallelism that is left unused when Cyclone
is used to replicate a single log. To exploit this parallelism we use Cyclone to
replicate multiple logs, each using a different instance of the RAFT consensus
protocol. We show how to exploit the commutativity inherent in key-value store
interfaces to make use of this parallelism in Cyclone's design in
Section~\ref{sec:parallelism}.  A detailed evaluation of Cyclone's replication
is provided in the context of RocksDB - a popular persistent key value store -
in Section~\ref{sec:evaluation}. We then discuss related work before concluding.
